A Triple Origin for Twin Blue Stragglers in Compact Binaries

### 2018-12-10 apjl.tex
Add files via upload
Added figures and update text

@@

\begin{abstract}
In this Letter, we propose a potential formation mechanism for twin blue stragglers in compact binaries that involves mass transfer from an evolved outer tertiary companion on to the inner binary via a circumbinary disk.  We apply our hypothetical scenario to the observed double blue straggler system Binary 7782 in the old open cluster NGC 188, and show that its observed properties are easily and even naturally reproduced within the context our proposed model.  Within the context of the hypothetical formation mechanism proposed here, the presented work predicts the following properties for the post-mass transfer double BS tertiary:  (1)  For the outer tertiary orbit, the initial orbital period should lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$ 1100 days, assuming initial masses for the inner binary components of m$_{\rm 1} =$ 1.1 M$_{\odot}$ and m$_{\rm 2} =$ 0.9 M$_{\odot}$ and an outer tertiary mass of m$_{\rm 3} =$ 1.4 M$_{\odot}$. (2)  Larger final WD masses, and hence core masses for the donor at the time of mass transfer, should correspond to larger final outer orbital periods for the tertiary. (3)  For the inner binary, the rotational axes of the BSs should be aligned with each other and the orbital plane of the outer tertiary WD. (4)  The BSs in the inner binary should have roughly equal masses, independent of their initial masses.  This predicts that the initially lower mass MS star should accrete the most, and should hence be polluted more significantly by any accreted material.  This could be observable in the surface layers of a radiative star (i.e., He, C and O if the donor is an RGB star, and/or s-process elements if the donor is an AGB star). (5) Twin BSs in compact binaries formed from the proposed mechanism should be more frequent in younger clusters with ages $\lesssim$ 4-6 Gyr, since the donor will have  radiative envelope.

@@

Blue straggler (BS) stars appear brighter and bluer than the main-sequence turn-off (MSTO) in a cluster colour-magnitude diagram (CMD) \citep[e.g.][]{simunovic14,simunovic16}.  Two primary channels for BS formation have been proposed; mass transfer from an evolved donor on to a main-sequence star in a binary star system \citep[e.g.][]{mccrea64,knigge09,mathieu09,leigh11,geller11,geller12,gosnell14,gosnell15}, and direct stellar collisions involving main-sequence stars mediated via direct interactions involving binaries \citep[e.g.][]{hills75,shara97,leigh07,leigh11,leigh13,hypki13}.  Other possible, albeit related, formation mechanisms include mergers of compact MS-MS binaries, and mergers of the inner MS-MS binaries of hierarchical triple star systems induced by Lidov-Kozai oscillations coupled with tidal damping \citep[e.g.][]{perets09}.

In spite of these specific predictions for the expected properties of BSs formed from each of the above production mechanisms, there exist many BSs with observed properties that defy these simple scenarios.  For example, in the old open cluster M67, there lurks a candidate triple system that is posited to host \textit{two} BSs \citep{vandenberg01,sandquist03}.  The observations suggest that the outer tertiary is itself a BS, with a mass $\sim$ 1.7 M$_{\odot}$ and orbiting the inner binary with a period of 1188.5 days \citep{sandquist03}.  The inner binary has a period of only 1.068 days \citep{vandenberg01}, and hosts a BS of mass 2.52 M$_{\odot}$.  Thus, in order to form this system after convolving its observed properties with its host cluster age, we require at least five stars in order to conserve the total system mass \citep{leigh11}.  As explained in \citet{leigh11}, this is strongly indicative of a dynamical origin for the system, and a single direct interaction involving a binary and a triple that resulted in two separate collision events is the most probable explanation for its origin (i.e., a single interaction involving two multiples with two or more stars is a more likely scenario to produce this system than two back-to-back direct binary-binary interactions).

Even more curious, there exists in the old open cluster NGC 188 a double BS binary, called Binary 7782.  Specifically, \citet{mathieu09} observed a compact and mildly eccentric (i.e., e $\sim$ 0.1) binary star system with an orbital period of $\sim$ 10 days hosting \textit{two} blue stragglers.  During a given binary-binary interaction, the probability that not one but \textit{two} direct MS-MS collisions occur is less than a percent \citep{leonard89,leigh11,leigh12}.  Typically, such binaries are very wide with long orbital periods.  Thus, dynamically, it is very difficult to form a compact binary composed of two collision products during a single direct interaction in a star cluster.  So, how did Binary 7782 form?  MENTION BIMODAL P-E DISTRIBUTION, AND HOW THIS COULD BE SUGGESTIVE OF A TRIPLE ORIGIN FOR SOME OF THE SUBSET OF OBSERVED BSs.

In this Letter, we propose a potential formation channel for Binary 7782, and compact double BS binaries in general, which involves mass transfer from an outer tertiary companion on to an inner MS-MS binary.  In section~\ref{dyn}, we constrain the range of initial (i.e., pre-mass transfer) orbital parameters for a hypothetical outer tertiary companion, using a combination of dynamical and stellar evolution-based constraints.  In Section~\ref{sims} we present the numerical simulations used to study the mass transfer process in our hypothetical triple system, computed using the \texttt{AMUSE} software package, in order to study the evolution of the inner and outer orbital parameters during mass transfer.  We summarize and discuss the implications of our results for compact double BS binaries and, more generally, mass transfer in stellar triples in Section~\ref{discussion}.

@@

Consider a hierarchical triple system with component masses m$_{\rm 1}$ and m$_{\rm 2}$ for the inner binary, and mass m$_{\rm 3}$ for the outer tertiary companion.  The inner and outer binary orbital semi-major axes are denoted a$_{\rm in}$ and a$_{\rm out}$, respectively.  We assume circular orbits for both the inner and outer orbits, and co-planar triples only since this is what has been observed for low-mass tertiaries \citep[e.g.][]{moe18,tobin18}.  This initial configuration for our assumed formation scenario for Binary 7782, described below, is depicted in Figure~\ref{fig:fig1}.

@@

We consider a scenario where the outer tertiary companion is filling its Roche lobe and is hence transferring mass to the inner binary.  The mass transfer stream gathers at the circularization radius a$_{\rm c}$, and forms a circumbinary disk.  Using conservation of angular momentum, we can equate the specific angular momentum of the accreted mass at the inner Lagrangian point of the donor star to the final specific angular momentum of the accretion stream at the circularization radius about the inner binary:
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - r_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer tertiary companion, a$_{\rm c}$ is the semi-major axis of the orbit about the inner binary corresponding to the circularization radius and v$_{\rm orb,c}$ is the orbital velocity at a$_{\rm c}$.  The distance from the center of mass corresponding to the outer tertiary companion defined by the Roche lobe is given by \citep{eggleton83}: 
\begin{equation}
\label{eqn:roche}
R_{\rm L} = \frac{0.49q^{2/3}}{0.6q^{2/3} + {\rm ln}(1 + q^{1/3})}a_{\rm out},
\end{equation}
where the mass ratio q is defined as q $=$ m$_{\rm 3}$/(m$_{\rm 1} +$ m$_{\rm 2}$).  Combining Equation~\ref{eqn:roche} with Equation~\ref{eqn:specangmom1}, we can solve for the circularization radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we require that a$_{\rm in} <$ a$_{\rm c}$.

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane for Binary 7782.  We assume initial component masses of m$_{\rm 1} =$ 1.1 M$_{\rm \odot}$ and m$_{\rm 2} =$ 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3} =$ 1.4 M$_{\rm \odot}$ for the outer tertiary.  We compare the circularization radius to the semi-major axis of the inner binary, for which we require a$_{\rm c} >$ a$_{\rm in}$, after folding in all constraints from the requirements for dynamical stability, an outer tertiary that is Roche lobe-filling, and a dynamically hard outer tertiary orbit.  Note that the range of plotted orbital periods P$_{\rm in}$ corresponding to a contact state for the inner binary, assuming R$_{\rm 1} =$ R$_{\rm 2} =$ 1 R$_{\odot}$, lies outside the range of plotted values for P$_{\rm in}$, since it does not contribute significantly to constraining the outer orbital properties of a hypothesized outer WD tertiary.  The thick horizontal solid red line shows the allowed range of outer semi-major axes, after folding in all of the aforementioned criteria.  As is clear, this makes a relatively narrow prediction for the allowed ranges of outer tertiary orbits, namely 2.2 $\times$ 10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days, for our assumed final donor mass.

\begin{figure}[ht!]
\plotone{fig2.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane allowed for the hypothetical outer tertiary orbit of Binary 7782.  The solid diagonal black lines show the period corresponding to the circularization radius a$_{\rm c}$ for the mass transfer stream coming from the outer tertiary, for different values of the mass ratio, namely q $=$ 0.1, 0.7, 1, 2 and 10.  We assume initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}$ is computed for the outer tertiary according to our assumed mass ratio (with our fiducial case corresponding to q $=$ 0.7.  We assume completely conservative mass transfer for this exercise, and a final mass for the outer tertiary of 0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal black line shows a rough criterion for dynamical stability in the triple, approximately following \citet{mardling99} (i.e., a$_{\rm in} <$ 0.1a$_{\rm out}$ is required for long-term dynamical stability in equal-mass co-planar triples).  The vertical solid red lines show the outer orbital periods corresponding to the hard-soft boundary assuming central velocity dispersions of $\sigma =$ 1, 5 and 10 km s$^{-1}$.  The vertical dashed black lines show the maximum outer orbital period P$_{\rm out}$ for which the outer tertiary companion is Roche lobe-filling, assuming a stellar radius of R$_{\rm 3} =$ 200 R$_{\odot}$ (CHANGE, POSSIBLY USING MESA CALCULATIONS?!).  The horizontal dashed red line shows the observed orbital period for Binary 7782, using its observed orbital period and our assumed final inner companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2} =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal red line shows the parameter space for P$_{\rm out}$ allowed after considering all of the aforementioned criteria.
\label{fig:fig2}}
\end{figure}

@@

We use smoothed-particle hydrodynamics (SPH) simulations to compute the time evolution of the mass transfer process.  We adopt the same initial particle masses as in Figure~\ref{fig:fig2}, namely initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}$ is computed for the outer tertiary according to our assumed mass ratio, with our fiducial case corresponding to q $=$ 0.7.  The initial orbital eccentricities are set to zero.  The inner and outer binary orbital semi-major axes are set to 0.1 AU and 1 AU, respectively (CORRECT).

\subsection{Initial Conditions} \label{ICs}

In this section, we describe and justify our choice of initial conditions for both our analytic calculations and smoothed-particle hydrodynamics simulations.

We adopt initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}$ = 1.4 M$_{\rm \odot}$ for the outer tertiary.  This choice for the initial mass of the outer tertiary is critical since it ensures that the donor star during the mass transfer process will have a radiative envelope \citep[e.g.][]{maeder09}.  In turn, this ensures that the mass transfer will be maximally conservative, such that the accretion stream will be maximally stable, accreting at a stable and constant rate \citep[e.g.][]{iben91}.

@@

@@

@@

Finally, in future work, we intend to explore the implications of the mechanism proposed here for other types of stellar triples.  The results highlight that specific theoretical predictions can often be made.  For example, if two white dwarfs form the inner compact binary, than circumbinary accretion could push one of them above the Chandrasekhar mass limit.  Upon explosion, its WD companion would be ejected at roughly the orbital velocity at the time of explosion, producing a hypervelocity WD.  This predicts hot hypervelocity WDs with masses very close to the Chandrasekhar mass limit for this scenario.  Similarly, if the two WDs are replaced with two neutron stars (NSs), then one NS will likely accrete to surpass the critical supernova limit, upon which time it explodes to become a black hole and receives a strong natal kick.  The NS companion would have a very comparable mass to the newly formed black hole progenitor at the time of explosion, and could potentially be used to constrain the maximum NS mass.

@@

N.W.C.L. acknowledges support from a Kalbfleisch Fellowship at the American Museum of Natural History.

\begin{thebibliography}{}

%\bibitem[\protect\citeauthoryear{Anderson et al.}{2008}]{anderson08}
% Anderson J.,  Sarajedini A., Bedin L. R., King I. R., Piotto G.,
%  Reid I. N., Siegel M., Majewski S. R., Paust N. E. Q., Aparicio A.,
%  Milone A. P., Chaboyer B., Rosenberg A. 2008, AJ, 135, 2055
%\bibitem[\protect\citeauthoryear{Artymowicz et al.}{1993}]{artymowicz93} Artymowicz P., Lin D. N. C., Wampler E. J. 1993, ApJ, 409, 592 
%\bibitem[\protect\citeauthoryear{Bartos et al.}{2017}]{Bartos17} Bartos I., Kocsis B., Haiman Z. \& M\'{a}rka S., 2017, ApJ, 835, 165
%\bibitem[\protect\citeauthoryear{Bartko et al.}{2009}]{Bartko09} Bartko H., Martins F., Trippe S., Fritz T. K., Genzel R., Ott T., Eisenhauer F., Gillessen S., Paumard T., Alexander T., Dodds-Eden K., Gerhard O., Levin Y., Mascetti L., Nayakshin S., Perets H. B., Perrin G., Pfuhl O., Reid M. J., Rouan D., Sternberg A., Trippe S., 2009, ApJ, 697, 1741
%\bibitem[\protect\citeauthoryear{Bartko et al.}{2010}]{Bartko10} Bartko H., Martins F., Fritz T. K., Genzel R., Levin Y., Perets H. B., Paumard T., Nayakshin S., Gerhard O., Alexander T., Dodds-Eden K., Eisenhauer F., Gillessen S., Mascetti L., Ott T., Perrin G., Pfuhl O., Reid M. J., Rouan D., Zilka M., Sternberg A., 2010, ApJ, 708, 834
%\bibitem[\protect\citeauthoryear{Baruteau et al.}{2011}]{Baruteau11} Baruteau C., Cuadra J. \& Lin D.N.C., 2011, ApJ, 726, 28
%\bibitem[\protect\citeauthoryear{Bedin et al.}{2010}]{bedin10} Bedin L. R., Cassisi S., Castelli F., Piotto G., Anderson J., Salaris M., Momany Y., Pietrinferni A. 2010, MNRAS 357, 1038
%\bibitem[\protect\citeauthoryear{Bender et al.}{2005}]{bender05} Bender R., Kormendy J., Bower G., et al. 2005, ApJ, 631, 280
%\bibitem[\protect\citeauthoryear{Bruzual \& Charlot}{2003}]{bruzual03} Bruzual G., Charlot S. 2003, MNRAS, 344, 1000 
%\bibitem[\protect\citeauthoryear{Casassus et al.}{2013}]{casassus13} Casassus S., Hales A., de Gregorio I., Dent B., Belloche A., G\"usten R., M\'enard F., Hughes A. M., Wilner D., Salinas V. 2013, A\&A, 553, 64 
%\bibitem[\protect\citeauthoryear{Ceillier et al.}{2017}]{ceillier17} Ceillier T., Tayar J., Mathur S., Salabert D., Garc\'ia R. A., Stello D., Pinsonneault M. H., van Sanders J., Beck P. G., Bloemen S. 2017, A\& A, 605, 111 

>>>>

\newcommand{\MSun}{\mbox{M$_\odot$}}
\newcommand{\RSun}{\mbox{R$_\odot$}}
\newcommand{\LSun}{\mbox{L$_\odot$}}

\def\Simon#1{{\bf {\color{red}[#1 -- Simon]}}}
\def\simon#1{{\bf {\color{red}[#1 -- Simon]}}}
\def\del#1{{\bf {\sout{#1}}}}
\def\replace#1#2{{\bf {\sout{#1} $\rightarrow$ {\bf #2}}}}

@@

\begin{abstract}

We propose a formation mechanism for twin blue stragglers in compact
binaries that involves mass transfer from an evolved outer tertiary
companion on to the inner binary via a circumbinary disk.  We apply
our hypothetical scenario to the observed double blue straggler system
Binary 7782 in the old open cluster NGC 188, and show that its
observed properties are naturally reproduced within the context our
proposed model.  Within the context of the hypothetical formation
mechanism proposed here, the presented work predicts the following
properties for the post-mass transfer double BS tertiary: (1) For the
outer tertiary orbit, the initial orbital period should lie between
220 days $\lesssim$ P$_{\rm out}$ $\lesssim$ 1100 days, assuming
initial masses for the inner binary components of m$_{\rm 1} =$ 1.1
M$_{\odot}$ and m$_{\rm 2} =$ 0.9 M$_{\odot}$ and an outer tertiary
mass of m$_{\rm 3} =$ 1.4 M$_{\odot}$. (2) Larger final WD masses, and
hence core masses for the donor at the time of mass transfer, should
correspond to larger final outer orbital periods for the tertiary. (3)
For the inner binary, the rotational axes of the BSs should be aligned
with each other and the orbital plane of the outer tertiary WD. (4)
The BSs in the inner binary should have roughly equal masses,
independent of their initial masses.  This predicts that the initially
lower mass MS star should accrete the most, and should hence be
polluted more effectively by the accreted material.  This could be
observable in the surface layers of a radiative star (i.e., He, C and
O if the donor is an RGB star, and/or s-process elements if the donor
is an AGB star). (5) Twin BSs in compact binaries formed from the
proposed mechanism should be more frequent in younger clusters with
ages $\lesssim$ 4-6 Gyr, since the donor will have radiative envelope.

@@

Blue straggler (BS) stars are brighter and bluer than the
main-sequence turn-off (MSTO) in a cluster colour-magnitude diagram
(CMD) \citep[e.g.][]{1953AJ.....58...61S,simunovic14,simunovic16}.
Two primary channels for BS formation have been proposed; mass
transfer from an evolved donor on to a main-sequence star in a binary
star system
\citep[e.g.][]{mccrea64,997A&A...328..143P,knigge09,mathieu09,leigh11,geller11,geller12,gosnell14,gosnell15},
and direct stellar collisions involving main-sequence stars mediated
via direct interactions involving binaries
\citep[e.g.][]{hills75,1997A&A...328..130P,shara97,leigh07,leigh11,leigh13,hypki13,2018arXiv181100058P}.
Other possible, albeit related, formation mechanisms include mergers
of compact main-sequence binaries (ms, ms), and mergers of the inner
binaries of hierarchical triple star systems induced by Lidov-Kozai
oscillations coupled with tidal damping \citep[e.g.][]{perets09}.

In spite of these specific predictions for the expected properties of
BSs formed from each of the above production mechanisms, there exist
many BSs with observed properties that defy these simple scenarios.
\del{For example, in the old open cluster M67, there lurks a candidate
  triple system that is posited to host \textit{two} BSs
  \citep{vandenberg01,sandquist03}.  The observations suggest that the
  outer tertiary is itself a BS, with a mass $\sim$ 1.7 M$_{\odot}$
  and orbiting the inner binary with a period of 1188.5 days
  \citep{sandquist03}.  The inner binary has a period of only 1.068
  days \citep{vandenberg01}, and hosts a BS of mass 2.52 M$_{\odot}$.}
In order to form this system after convolving its observed properties
with its host cluster age, we require at least five stars in order to
conserve the total system mass \citep{leigh11}.  As explained in
\citet{leigh11}, this is strongly indicative of a dynamical origin for
the system, and a single direct interaction involving a binary and a
triple that resulted in two separate collision events is the most
probable explanation for its origin (i.e., a single interaction
involving two multiples with two or more stars is a more likely
scenario to produce this system than two back-to-back direct
binary-binary interactions) \citep{2004MNRAS.350..615G}.

Even more curious, there exists in the old open cluster NGC 188 a
double BS binary, called Binary 7782.  Specifically, \citet{mathieu09}
observed a compact and mildly eccentric (i.e., e $\sim$ 0.1) binary
star system with an orbital period of $\sim$ 10 days hosting two blue
stragglers.  During a given binary-binary interaction, the probability
that not one but two direct (ms, ms) collisions occur is less than
$10^{-2}$ \citep{leonard89,leigh11,leigh12}.  Typically, such binaries
have long orbital periods. Dynamically, it difficult to form a
short-period binary composed of two collision products during a
collision interaction in a star cluster \citep{2011Sci...334.1380F}.
So, how did Binary 7782 form?  MENTION BIMODAL P-E DISTRIBUTION, AND
HOW THIS COULD BE SUGGESTIVE OF A TRIPLE ORIGIN FOR SOME OF THE SUBSET
OF OBSERVED BSs.

We propose a formation channel for Binary 7782, and compact double BS
binaries in general, which involves mass transfer from an outer
tertiary companion on to an inner MS-MS binary.  In section~\ref{dyn},
we constrain the range of initial (i.e., pre-mass transfer) orbital
parameters for a hypothetical outer tertiary companion, using a
combination of dynamical and stellar evolution-based constraints.  In
Section~\ref{sims} we present the numerical simulations used to study
the mass transfer process in our hypothetical triple system, computed
using the \texttt{AMUSE}\cite{AMUSE} software package, in order to
study the evolution of the inner and outer orbital parameters during
mass transfer.  We summarize and discuss the implications of our
results for compact double BS binaries and, more generally, mass
transfer in stellar triples in Section~\ref{discussion}.

@@

Consider a hierarchical triple system with component masses m$_{\rm
  1}$ and m$_{\rm 2}$ for the inner binary, and mass m$_{\rm 3}$ for
the outer tertiary companion.  The inner and outer binary orbital
semi-major axes are denoted a$_{\rm in}$ and a$_{\rm out}$,
respectively.  We assume circular orbits for both the inner and outer
orbits, and co-planar triples only since this is what has been
observed for low-mass tertiaries \cite{2010yCat..73890925T} \simon{I
  think that {\em moe18} and {\em tobin18} are not the right
  references}.  This initial configuration for our assumed formation
scenario for Binary 7782, described below, is depicted in
Figure~\ref{fig:fig1}.

@@

We consider a scenario where the outer star is filling its Roche lobe
and is transfers mass to the inner binary.  The mass transfer stream
gathers at the circularization radius a$_{\rm c}$, and forms a disk
around the inner binary.  Using conservation of angular momentum, we
equate the specific angular momentum of the accreted mass at the inner
Lagrangian point of the (outer) donor star to the final specific
angular momentum of the accretion stream at the circularization radius
about the inner binary, this results in
\simon{replaced lower case $r$ with upper case.}
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - R_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer
tertiary companion, a$_{\rm c}$ is the semi-major axis of the orbit
about the inner binary corresponding to the circularization radius and
v$_{\rm orb,c}$ is the orbital velocity at a$_{\rm c}$.  The distance
from the center of mass corresponding to the outer tertiary companion
defined by the Roche lobe is given by \citep{eggleton83}.
\simon{Do we really need this eq?}
\begin{equation}
\label{eqn:roche}
R_{\rm L} = \frac{0.49q^{2/3}}{0.6q^{2/3} + {\rm ln}(1 + q^{1/3})}a_{\rm out},
\end{equation}
where the mass ratio q is defined as q $=$ m$_{\rm 3}$/(m$_{\rm 1} +$
m$_{\rm 2}$).  Combining Equation~\ref{eqn:roche} with
Equation~\ref{eqn:specangmom1}, we can solve for the circularization
radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we
require that a$_{\rm in} <$ a$_{\rm c}$.

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm
  out}$-P$_{\rm in}$-plane for Binary 7782.  We assume initial
component masses of m$_{\rm 1} =$ 1.1 M$_{\rm \odot}$ and m$_{\rm 2}
=$ 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}
=$ 1.4 M$_{\rm \odot}$ for the outer tertiary.  We compare the
circularization radius to the semi-major axis of the inner binary, for
which we require a$_{\rm c} >$ a$_{\rm in}$, \del{after folding in all
  constraints from the requirements for dynamical stability}\simon{can
  we not simply state the constraints?}, an outer tertiary that is
Roche lobe-filling, and a dynamically hard outer tertiary orbit.  Note
that the range of plotted orbital periods P$_{\rm in}$ corresponding
to a contact state for the inner binary, assuming \simon{This looks
  weird:} R$_{\rm 1} =$ R$_{\rm 2} =$ 1 R$_{\odot}$, lies outside the
range of plotted values for P$_{\rm in}$, since it does not contribute
significantly to constraining the outer orbital properties of a
hypothesized outer WD tertiary.  The thick horizontal solid red line
shows the allowed range of outer semi-major axes, after folding in all
of the aforementioned criteria.  As is clear, this makes a relatively
narrow prediction for the allowed ranges of outer tertiary orbits,
namely 2.2 $\times$ 10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1
$\times$ 10$^3$ days, for our assumed final donor mass.

\begin{figure}[ht!]
\plotone{fig2.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane
  allowed for the hypothetical outer tertiary orbit of Binary 7782.
  The solid diagonal black lines show the period corresponding to the
  circularization radius a$_{\rm c}$ for the mass transfer stream
  coming from the outer star, for different values of the mass ratio,
  namely q $=$ 0.1, 0.7, 1, 2 and 10 \simon{before or after mass
    transfer?}.  \simon{what is the purpose of extending q to 0.1 and
    10?} We assume initial component masses of m$_{\rm 1}$ = 1.1
  M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner
  binary components, and m$_{\rm 3}$ is computed for the outer
  tertiary according to our assumed mass ratio (with our fiducial case
  corresponding to q $=$ 0.7.  We assume completely conservative mass
  transfer for this exercise, and a final mass for the outer tertiary
  of 0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal
  black line shows a rough criterion for dynamical stability in the
  triple, approximately following \citet{mardling99} (i.e., a$_{\rm
    in} <$ 0.1a$_{\rm out}$ is required for long-term dynamical
  stability in equal-mass co-planar triples).  The vertical solid red
  lines show the outer orbital periods corresponding to the hard-soft
  boundary assuming central velocity dispersions of $\sigma =$ 1, 5
  and 10 km s$^{-1}$.  The vertical dashed black lines show the
  maximum outer orbital period P$_{\rm out}$ for which the outer
  tertiary companion is Roche lobe-filling, assuming a stellar radius
  of R$_{\rm 3} =$ 200 R$_{\odot}$ (CHANGE, POSSIBLY USING MESA
  CALCULATIONS?!).  The horizontal dashed red line shows the observed
  orbital period for Binary 7782, using its observed orbital period
  and our assumed final inner companion masses (i.e., m$_{\rm 1} =$
  m$_{\rm 2} =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal
  red line shows the parameter space for P$_{\rm out}$ allowed after
  considering all of the aforementioned criteria.  \simon{I think that
    the figure needs work. I like putting the parameter space in terms
    of orbital periods, but the q=0.1 and q=10 constraints seems
    not very practical. The red labels are not readable, and I am not
    really sure how to read the figure. What about the alternative in
    the next figure, which shows tertiary mass vs. orbital period?}
\label{fig:fig2}}
\end{figure}

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_minimumstablesize.pdf}
\caption{Orbital separation as a function of the mass of the
  Roche-lobe filling outer star. The red curves gives the size at the
  first giant branch (thick curve) at at the AGB (thin curve). The
  blue curve the orbital separation at the moment this giant fills its
  Roche-lobe assuming a circular orbit and a total inner-binary mass
  of 2.0\,\MSun. The green curves give, under these conditions, the
  maximum orbital separation for a stable inner binary.
\label{fig:tertiarymass_vs_size}}
\end{figure}

In Fig.\,\ref{fig:tertiarymass_vs_size} we present..

@@

We perform simulations of a triple star system of which the outer star
overfills its Roche lobe while the inner binary remains detached. The
calculations start by evolving the three stars to the same age, which
is selected such that the outer-most star fills its Roche lobe.
Constraints for the initial conditions are derived in the previous
\S\, but in \S\,\ref{Sect:ICs} we summarize the adopted initial
conditions for the simulations.

RLOF in triples using a coupled integrator to follow the complex
hydrodynamics of the mass transfer from the Roche-lobe filling outer
star to the inner binary, while keeping track of the gravitational
dynamics of the stars.  The equations of motion of the inner binary is
solved using the semi-symplectic direct N-body integrator
\texttt{Huayno} \citep{2012NewA...17..711P}. The hydrodynamics is
performed with the smoothed-particles hydrodynamics code
\texttt{Gadget2} \citep{2000ascl.soft03001S}, using an adiabatic
equation of state.  The two inner binary stars are treated as point
masses, but we allow them to accrete gas from the outer star.  This is
realized using sink-particles that co-move with the mass points in the
gravity code. While the inner two stars accrete mass, they also
accrete the appropriate amoun of angular momentum from the gas.  and
the the N-body integrator accounts of this appropriately.  For the
radius of the sink particles we adopt $10 R_\odot$ for both stars.

The N-body code as well as the hydrodynamics solver operate using
their own internal time-steps. The coupling between the two codes is
realized using the \texttt{Bridge} method in the AMUSE framework
\citep[see Sect.\.4.3.1 in][]{2013CoPhC.183..456P}.  This coupled
integrator is based on the splitting of the Hamiltonian much in the
same way as is done with two different gravity solvers by
\cite{2007PASJ...59.1095F}. With the adopted scheme, the
hydrodynamical solver is affected by the gravitational potential of
its own particles, as well as the gravitational potential of the inner
binary. The hydrodynamics, in particular the gas drag, in turn affects
the orbits of the two inner stars. With \texttt{Bridge} we realize a
second order coupling between the gravitational dynamics and the
hydrodynamics.  The interval at which the gravity and hydrodynamics
interact via \texttt{Bridge} depends on the parameters of the system
we study, but typically we achieve converged solutions when this time
step is about a fraction of 1/100 of the inner binary orbital period.

\subsection{Initial Conditions} \label{Sect:ICs}

We adopt initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$
and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components,
and m$_{\rm 3}$ = 1.2 M$_{\rm \odot}$ for the tertiary star.  We
evolved the 1.2\,\MSun\, star using the MESA stellar evolution code
\cite{2011ApJS..192....3P} to a radius of about 255\,\RSun.  By this
time the star has lost 0.13\,\MSun\, in a stellar wind.  With these
parmeters the orbital separation of the outer binary becomes 3.86\,au,
and the mass loss has driven a slight eccentricity from initiall
circular orbit to about 0.0082 at the moment of Roche-lobe
overflow. We chose a relative initial inclination of $i = 9^\circ$
between the inner and the outer orbit.

In fig.\,\ref{fig:topview_at_t0} we present a top view of the triple
system at the onset of Roche-lobe overlofw.

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_t0_N80000_M012MSun1109MSun_a02au_e00_inc9deg.pdf}
\caption{top view of the triple system in which the outer star
  (represented by 80000\,sph particles and a core particle of
  0.4\,\MSun (blue bullet). The two companion stars are represented as
  red bullets (to the right).
\label{fig:topview_at_t0}}
\end{figure}

@@

In fig.\,\ref{fig:topview_at_t1000day} we show the same triple at
after 1000\,days of evolution.

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_t600_N80000_M012MSun1109MSun_a02au_e00_inc9deg.pdf}
\caption{Same as Fig.\,\ref{fig:topview_at_t0} but at an age of $t
  \simeq 996days$.
\label{fig:topview_at_t1000day}}
\end{figure}

In fig.\,\ref{fig:mass_evolution_for_M1.2M1.1M0.9a02e00i09}

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_MvsTime_N80k_M012R100RSun1109MSun_a02au.pdf}
\caption{Evolution of the masses (normalized to those at the moment of
  Roche-lobe overflow) of each of the three stars as a function of
  time for the first 1000\,days. (see Fig.\.\ref{fig:topview_at_t0}
  for a face-on representation of the initial conditions and
  fig.\,\ref{fig:topview_at_t1000day} for the final conditions}.  The
green and orange dashed curves give the masses of the inner two stars,
whereas the solid blue curve is the mass of the outer most star.
\label{fig:mass_evolution_for_M1.2M1.1M0.9a02e00i09}
\end{figure}

@@

This
choice for the initial mass of the outer tertiary is critical since it
ensures that the donor star during the mass transfer process will have
a radiative envelope \citep[e.g.][]{maeder09}.  In turn, this ensures
that the mass transfer will be maximally conservative, such that the
accretion stream will be maximally stable, accreting at a stable and
constant rate \citep[e.g.][]{iben91}.

@@

Finally, in future work, we intend to explore the implications of the mechanism proposed here for other types of stellar triples.  The results highlight that specific theoretical predictions can often be made.  For example, if two white dwarfs form the inner compact binary, than circumbinary accretion could push one of them above the Chandrasekhar mass limit.  Upon explosion, its WD companion would be ejected at roughly the orbital velocity at the time of explosion, producing a hypervelocity WD.  This predicts hot hypervelocity WDs with masses very close to the Chandrasekhar mass limit for this scenario.  Similarly, if the two WDs are replaced with two neutron stars (NSs), then one NS will likely accrete to surpass the critical supernova limit, upon which time it explodes to become a black hole and receives a strong natal kick.  The NS companion would have a very comparable mass to the newly formed black hole progenitor at the time of explosion, and could potentially be used to constrain the maximum NS mass.  More work needs to be done here to evaluate whether or not there is any parameter space for the products of this scenario that will allow for a mass determination for the original NS companion in the putative compact NS-NS binary.

@@

N.W.C.L. acknowledges support from a Kalbfleisch Fellowship at the
American Museum of Natural History.  SPZ would like to thank Norm
Murray and CITA for their hospitality during my long-term visit.

\input /home/spz/Latex/lib/bib/references

%\bibliographystyle{/home/spz/Latex/lib/styles/elsevier/elsarticle-num} 
%\bibliography{references}
